Are we living in a depressive society?

### 2024-04-14 outputs/paper/paper.qmd
Update paper.qmd
update paper

---
title: "Are we living in a depressive society?"
subtitle: "Trends in symptoms of depressive disorder in the US population from May 2020 to March 2024"
author: Chay Park
thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Anxiety_depressive_disorder."
date: today
date-format: long
abstract: "This study explores how socio-cultural factors like age, sex, race/ethnicity, and education level influenced depressive disorder in the U.S. during and after the COVID-19 pandemic, analyzing data from 2020 to 2024. Using advanced statistical methods in R, it uncovered significant variations in mental health impacts across different demographic groups. The research reveals a complex relationship between socio-cultural factors and mental health, highlighting the increased vulnerability of certain populations during global crises. The findings contribute to the development of targeted public health measures to alleviate mental health issues amid and following pandemics, enhancing our understanding of the impact of socio-cultural factors on mental well-being in times of crisis."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This paper addresses this gap by analyzing data from the U.S. Census Bureau's Household Pulse Survey, initiated in response to the pandemic to assess its effects on American households, including mental wellness (U.S. Census Bureau, 2020). Our analysis focuses on the correlation between the prevalence of anxiety or depressive disorders and socio-cultural factors such as age, sex, race/ethnicity, and education level during the pandemic and post pandemic years of 2020 to 2024. Specifically, the estimand of this study is the difference in the rates of symptoms of anxiety or depressive disorders across individuals, correlated with socio-cultural factors including age, sex, race/ethnicity, and education level.

@@

* How do adepressive rates differ based on education level?

@@

# Plotting the line graph
p <- ggplot(filtered_data, aes(x = Date, y = Value, group = Indicator, color = Indicator)) +
  geom_line() +
  geom_hline(yintercept = seq(10, 45, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL,
       y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.direction = "vertical",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

@@

The @fig-andtrend-graph overall trend in US anxiety and depressive disorder rates fluctuate over time, with the highest rates recorded in the end of 2020 and the beginning of 2021 at an average rate above 40% symptom rates for both anxiety disorder or depressive disorder. Subsequently, symptoms of anxiety disorder showing similar trends with about 10% gap in-between. Lastly, the symptoms of depressive disorder followed by the anxiety disorder records the lowest out of three indicators, with an average of 23% symptom rates over the period.

The overall trend in US anxiety and depressive disorder rates shows significant fluctuations, with several notable peaks indicating heightened rates of reported symptoms. For instance, one of the highest peaks occurs in early 2021, where the rate of symptoms for combined anxiety and depressive disorders reaches approximately 45%. This peak reflects a substantial increase compared to earlier periods. Another significant peak, observed in mid-2022, sees rates approaching closer to 36-37%, illustrating another period of notable increase in symptoms among the population.

These peaks are interspersed with valleys, such as in late 2020, where rates temporarily dip to around 25%, showing the episodic nature of these conditions. The plot also highlights a general trend of increasing rates from the onset of the data in 2020 through to peaks in 2021, followed by some fluctuation but maintaining relatively high levels overall.

The data indicates that the combined symptom rates of both anxiety and depressive disorders generally present higher values, consistently staying above 30% after the initial 2020 data points. This trend underscores the overlapping nature of these mental health issues and their persistent impact on the US population.

The graph, depicted as @fig-andtrend-graph, provides a line graph visualization of these rates over time. This visual representation allows for clear observation of trends and seasonal variations, with specific dates marked on the x-axis to aid in pinpointing changes. For example, a significant increase is noted as we move from late 2020 into early 2021, correlating with various global and national events that may have influenced mental health.

@@

```
@fig-ageandtrend-graph illustrates the trends in depressive disorder rates by population subgroup segmented by age, covering the period from May 2020 to March 2024. This graph reveals several key insights about how depressive symptoms have fluctuated across different age groups during and potentially influenced by the ongoing global and societal changes.

### Analysis of Trends by Age Group
The line graph distinguishes each age group with a separate line, showing the trend of depressive disorder rates for each demographic:

@@

## Trends in Depressive Disorder Rates by Population Subgroup by Sex

@@

## Trends in Depressive Disorder Rates by Population Subgroup by Race and Ethnicity

@@

## Trends in Depressive Disorder Rates by Population Subgroup by Education

>>>>

---
title: "Are we living in a depressive society?"
subtitle: "Trends in symptoms of depressive disorder in the US population from May 2020 to March 2024"
author: Chay Park
thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Anxiety_depressive_disorder."
date: today
date-format: long
abstract: "This study explores how socio-cultural factors like age, sex, race/ethnicity, and education level influenced depressive disorder in the U.S. during and after the COVID-19 pandemic, analyzing data from 2020 to 2024. Using statistical methods in R, it uncovered variations in mental health impacts across different demographic groups. The research reveals a complex relationship between socio-cultural factors and mental health, highlighting the increased vulnerability of certain populations during global crises. The findings contribute to the development of targeted public health measures to alleviate mental health issues amid and following pandemics, enhancing our understanding of the impact of socio-cultural factors on mental well-being in times of crisis."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

This paper addresses this gap by analyzing data from the U.S. Census Bureau's Household Pulse Survey, initiated in response to the pandemic to assess its effects on American households, including mental wellness (U.S. Census Bureau, 2020). Our analysis focuses on the correlation between the prevalence of anxiety or depressive disorders and socio-cultural factors such as age, sex, race/ethnicity, and education level during the pandemic and post pandemic years of 2020 to 2024. Specifically, the estimand of this study is the difference in the rates of symptom of depressive disorders across individuals, correlated with socio-cultural factors including age, sex, race/ethnicity, and education level.

@@

* How do depressive rates differ based on education level?

@@

# Plotting the line graph
p <- ggplot(filtered_data, aes(x = Date, y = Value, group = Indicator, color = Indicator)) +
  geom_line() +
  geom_hline(yintercept = seq(10, 45, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL,
       y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.direction = "vertical",
        legend.text = element_text(family = "Times"),
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

@@

The overall trend in US anxiety and depressive disorder rates @fig-andtrend-graph shows significant fluctuations over the observed period, with several notable peaks indicating heightened rates of reported symptoms. For instance, one of the highest peaks occurs in early 2021, where the rate of symptoms for combined anxiety and depressive disorders reaches approximately 45%. This peak reflects a substantial increase compared to earlier peeriods. Another significant peak, observed in mid-2022, sees rates approaching closer to 36-37%, illustrating another period of notable increases in symptoms among the population.

These peaks are interspersed with valleys, such as in late 2020, where rates temporarily din to around 25%, demonstrating the episodic nature of these conditions. The plot also highlights a general trend of increasing rates from the onset of the data in 2020 through to peaks in 2021, followed by some fluctutation but maintaining relatively high levels overall.

In late 2023, there is a notable decline in the symptom rates for both anxiety and depressive disorders, suggesting a potential easing of these mental health challenges or the effectiveness of interventions and strategies implemented over the period. This decrease is significant as it may indicate periods of relief amidst the prevailing trends of increasing disorder rates, which could be crucial for future public health planning and policymaking.

The @fig-andtrend-graph provides a line graph visualization of these rates over time. This visual representation allows for clear observation of trends and seasonal variations, with specific dates marked on the x-axis to aid in pinpointing changes.

@@

```
@fig-ageandtrend-graph illustrates the trends in depressive disorder rates by population subgroup segmented by age, covering the period from May 2020 to March 2024. This graph reveals several key insights about how depressive symptoms have fluctuated across different age groups during and potentially influenced by the ongoing global and societal changes. The line graph distinguishes each age group with a separate line, showing the trend of depressive disorder rates for each demographic:

@@

@@

### Trends in Depressive Disorder Rates by Population Subgroup by Race and Ethnicity

@@

### Trends in Depressive Disorder Rates by Population Subgroup by Education
